\documentclass{school-22.211-notes}
\date{March 19, 2012}

\begin{document}
\maketitle

\lecture{One-Group Diffusion: Analytical Solutions, Part 1}
\topic{Helmholtz Equation}
We start with the one-group diffusion equation,
\eqn{ -\divergence D(\vecr) \gradient \phi(\vecr) + \Sigma_a(\vecr) \phi(\vecr) = \frac{1}{\keff} \nu \Sigma_f(\vecr) \phi(\vecr) }
Assume homogeneous material, that is, spatial constant cross section,
\eqn{ -D \laplacian \phi(\vecr) + \Sigma_a \phi(\vecr) = \frac{1}{\keff} \nu \Sigma_f \phi(\vecr) }
Rearranging and defining a buckling term, 
\eqn{ \laplacian \phi(\vecr) + \underbrace{\frac{ \frac{\nu \Sigma_f}{\keff} - \Sigma_a}{D} }_{B^2} \phi(\vecr) = 0 }
We obtain a classic Helmholtz Equation,
\eqn{ \boxed{\laplacian \phi(\vecr) + B_m^2 \phi(\vecr) = 0 } }
Helmholtz Equation implies that,
\eqn{ B_m^2 = - \frac{\laplacian \phi(\vecr)}{\phi(\vecr)} }
Notice,
\begin{enumerate}
\item Since $B_m^2$ is a constant, that is to say $\frac{\laplacian \phi(\vecr)}{\phi(\vecr)}$ is a constant, that is to say $\phi(\vecr)$ has a constant curvature. 
\item Material buckling: 
  \eqn{ B_m^2 = \frac{\frac{\nu \Sigma_f}{\keff} - \Sigma_a}{D} }
  which matches geometry buckling, 
  \eqn{ \keff = \frac{\nu \Sigma_f}{DB_g^2 + \Sigma_a} }
  where $DB^2_g$ is the leakage per unit volume per unit flux. Notice $\keff$ does not depend on volume or flux. 

\item Critical buckling: when a system is critical, the material buckling is uniquely determined by cross sections, 
  \eqn{ B_m^2 = \frac{\nu \Sigma_f - \Sigma_a}{D} = \frac{\frac{\nu \Sigma_f}{\Sigma_a} - 1}{\frac{D}{\Sigma_a}} = \frac{\kinf - 1}{M^2} }
  \hi{Migration area} $M^2 = \frac{D}{\Sigma_a}$ is a measurement of the amount of travelling before absorption. 


\item Solutions exist only for certain values of the buckling such that the flux is everywhere positive and vanishing on outer (or extrapolated) surfaces; we define these unique values as `geometrical buckling' $B_g^2$. The allowable values of geometrical buckling that satisfy the boundary conditions are uniquely determined reactor geometry.

\item For a criticality problem, we can solve find the $B^2$ such that the system is critical. Hence giving geometry, only certain materials would make the system critical; given material, only certain geometries would make the system critical. That is, given a reactor, it would be critical if and only if the material and the geometry satisfy,
  \eqn{ B_g^2 = B_m^2} 

\item Solution only exists when $\kinf > 1$ (exception: if there is external source, $\kinf$ can be less than 1 and the system is still critical). 


\end{enumerate}



Talbe~\ref{simple-geometry-laplacian} lists a couple of one group homogeneous geometry Laplacians. 
\begin{table}
  \centering
  \begin{tabular}{|l|l|} \hline
    Slab & $\dphidxn2 + B^2 \phi(x) = 0$ \\ \hline
    Sphere & $\dphidrn2 + \frac{2}{r} \dphidr + B^2 \phi(r) = 0$ \\ \hline
    Infinite Cylinder & $\dphidrn2 + \frac{1}{r} \dphidr + B^2 \phi(r) = 0$ \\ \hline
    Finite Cylinder & $\dphidrn2 + \frac{1}{r} \dphidr + \dphidzn2 + B^2 \phi(r,z) = 0$ \\ \hline
    Cartesian & $\dphidxn2 + \dphidyn2 + \dphidzn2 + B^2 \phi(x,y,z) = 0$ \\ \hline
  \end{tabular}
\caption{Simple Geometry Laplacians} \label{simple-geometry-laplacian}
\end{table}




\clearpage
\topic{One Group Criticality Problem}
 Slab $\in  \left[- \frac{L_0}{2}, \frac{L_0}{2} \right]$:
  \eqn{ \dphidxn2 + B^2 \phi (x) &= 0, &\phi(x) &= A \cos (Bx) + C \sin(Bx) }
BCs: $\phi(\pm L/2) = 0$, $\frac{L}{2} = \frac{L_0}{2} + 0.711 \lambda_{tr}$. Two equations two unknowns, 
  \eqn{ \left[ \begin{array}{cc} \cos(BL/2) & \sin(BL/2) \\ \cos(BL/2) & -\sin(BL/2) \end{array} \right] \left[ \begin{array}{cc} A \\ C \end{array} \right] = 0 }
  Set the determinant to be zero, we get $-2 \cos (BL/2) \sin (BL/2) = 0$. There are two possibilities: 
\eqn{ B_n&= \frac{n\pi}{L}  & \phi(x) &= \left\{ 
  \begin{array}{cc} 
    A_n \cos (B_n x) & n=1,3,5, \cdots \\
    A_n \sin (B_n x) & n=2,4,6, \cdots 
  \end{array} \right. }
But in order for $\phi(x) \ge 0$ everywhere, only $n=1$ is possible; that is, 
\eqn{ \phi(x) = A \cos \frac{\pi x}{L} }
and the criticality condition implies that, 
\eqn{ \frac{\nu \Sigma_f - \Sigma_a}{D}  = \left( \frac{\pi}{L} \right)^2 }

\begin{table}
  \small
  \makebox[\textwidth][c]{
  \begin{tabular}{|l|l|l|l|} \hline
    Geometry & Diffusion Equation & Flux & Geometrical Buckling \\ \hline
    Slab $\in  \left[- \frac{L}{2}, \frac{L}{2} \right]$ &$\dphidxn2 + B^2 \phi = 0$ & $\phi = A \cos \left( \frac{\pi x}{L} \right)$ & $B^2 = \left( \frac{\pi}{L} \right)^2$ \\ 
    Sphere $\in [0, R]$  & $\dphidrn2 + \frac{2}{r} \dphidr + B^2 \phi = 0$& $\phi= \frac{A}{r} \sin \left( \frac{\pi r}{R} \right)$ & $B^2 = \left( \frac{\pi}{R} \right)^2$ \\
    Cylinder ($z: \infty$) & $\dphidrn2 + \frac{1}{r} \dphidr + B^2 \phi = 0$ & $\phi = A J_0 \left( \frac{2.405 r}{R} \right)$ & $B^2 = \left( \frac{2.405}{R} \right)^2$ \\ 
    Cylinder (z: $\pm H/2$) & $\dphidrn2 + \frac{1}{r} \dphidr + \dphidzn2 + B^2 \phi = 0$ & $\phi = A J_0\left( \frac{2.405 r}{R} \right) \cos \left( \frac{\pi z}{H} \right)$ & $B^2 = \left( \frac{2.405}{R} \right)^2 + \left( \frac{\pi}{H} \right)^2$ \\ 
    Parallelepiped ($\pm \frac{L_i}{2}$) & $\dphidxn2 + \dphidyn2 + \dphidzn2 + B^2 \phi = 0$ & $\phi = A \cos \left( \frac{\pi x}{L_x} \right) \cos \left( \frac{\pi y}{L_y} \right) \cos \left( \frac{\pi x}{L_z} \right)$ & $B^2 = \left( \frac{\pi}{L_x} \right)^2 + \left( \frac{\pi}{L_y} \right)^2 + \left( \frac{\pi}{L_z} \right)^2$ \\ \hline
  \end{tabular}
}
  \caption{One Group Fundamental Mode Eigenvalues and Eigenvectors} \label{eigen-values}
\end{table}
\normalsize

Observations:
\begin{itemize}
\item The lowest node is the only one remains after the source is gone. 
\item For any positive value of materials buckling, there is a unique critical size for each reactor geometry. 
\item Example: know how to find height-to-diameter to minimize leakage. The optimal cylinder has a H/D of around 0.92. 
\end{itemize}


\clearpage
\topic{One Group Source Problem} \label{one-group-source-problem-subcritical}
Consider a subcritical multiplying medium, slab geometry from $-L/2$ to $L/2$, and with a source. 
\begin{align}
  - D \dphidxn2 + \Sigma_a \phi(x) &= \nu \Sigma_f \phi(x) + S(x) \\
  \dphidxn2 + B_m^2 \phi(x) &= -\frac{S(x)}{D} 
\end{align}
Because $B^2 = \frac{\nu \Sigma_f - \Sigma_a}{D} < 0 $ for a subcritical condition, we can re-write $B^2 = - |B|^2$,
\eqn{ \dphidxn2 - |B|^2 \phi(x) = \tilde{S}(x) }
The general/homogeneous solution is,
\eqn{ \phi_H (x) = A e^{|B|x} + Ce^{-|B|x} = A \cosh (|B|x) + C \sinh (|B| x) }
The particular solution depends on source $\tilde{S}(x)$. Apply BCs $\phi \left( \pm \frac{L}{2} \right) = 0$, we get the boundary conditions in the matrix form,
\eqn{ \left[ \begin{array}{cc} 
\cosh(|B|L/2) & \sinh(|B|L/2) \\
\sinh(|B|L/2) & \cosh(|B|L/2) 
\end{array} \right] 
\left[ \begin{array}{c}
A \\ C \end{array} \right] 
= - \left[ \begin{array}{c} 
\phi_p(L/2) \\ \phi_p(-L/2) \end{array} \right] }
Interpretations:
\begin{enumerate}
\item Coefficients A and C are uniquely determined for given source distribution; 
\item There always exists a physically realizable solution (no critical buckling!);
\item In the limit of $S\to 0$, the only physical solution is the trivial solution (\textcolor{red}{shouldn't the only solution be the general solution?}). 
\end{enumerate}

\end{document}
